---
title: "Bootstrapping Simulation Study"
author: "Jimmy Kelliher"
date: "2/8/2022"
output: html_document
---

```{r setup, include=FALSE}
# load necessary packages
library(tidyverse)

# set knitr defaults
knitr::opts_chunk$set(
    echo      = TRUE
  , fig.align = "center"
  , fig.width = 6
  , fig.asp   = .6
  , out.width = "90%"
)

# set theme defaults
theme_set(
  theme_bw() +
  theme(
    legend.position = "bottom"
    , plot.title    = element_text(hjust = 0.5)
    , plot.subtitle = element_text(hjust = 0.5)
    , plot.caption  = element_text(hjust = 0.0)
  )
)

# set color scale defaults
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill   = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete   = scale_fill_viridis_d
```

Let the logit function $h : (0, 1) \to \mathbb{R}$ be given by

  \[ h(x) = \log \left( \frac{x}{1 - x} \right). \]

Fix $n \in \mathbb{N}$, and consider a population of individuals $i \in \{ 1, \ldots, n \}$. Let $X_{1i}, X_{2i}, X_{3i} \sim N(0, 1)$ be covariates associated with individual $i$. Let $T_i$ be a binary variable indicating some treatment, and let the data-generating process for $T_i$ satisfy

  \[
    h \Big( P \big(T_i = 1 \; | \; \mathbf{X}_i, \boldsymbol{\alpha} \big) \Big)
    = \alpha_0 + \alpha_1 X_{1i} + \alpha_2 X_{2i} + \varepsilon_i
  \]

for $\varepsilon_i \sim N(0, \sigma_\varepsilon^2)$. That is, the probability of receiving treatment $T_i$ for individual $i$ is determined by $X_{1i}, X_{2i},$ and random noise. Let $Y_i$ be some outcome variable with data-generating process

  \[
    Y_i
    = \beta_0 + \beta_{1i} T_i + \beta_2 X_{2i} + \beta_3 X_{3i} + \eta_i
  \]

for $\eta_i \sim N(0, \sigma_\eta^2)$. That is, the outcome for individual $i$ is determined by $T_i, X_{2i}, X_{3i},$ and random noise, as well as individual treatment effect $\beta_{1i} \sim N(\beta_1, \sigma^2)$.


```{r}
set.seed(8160)

# set parameters
n <- 10^3

alpha_0 <- -1.00
alpha_1 <-  0.25
alpha_2 <-  4.00

beta_0  <- 0
beta_1  <- 4
beta_2  <- 2
beta_3  <- 3

sigma_t <- 1 # standard deviation of noise in treatment DGP
sigma_y <- 1 # standard deviation of noise in outcome DGP
sigma_b <- 1 # standard deviation of distribution of individual treatment effects

# define the distribution of individual treatment effects
beta_1i <- rnorm(n, beta_1, sigma_b^2)
hist(beta_1i)

# define covariates
x_1 <- rnorm(n, 0, 1)
x_2 <- rnorm(n, 0, 1)
x_3 <- rnorm(n, 0, 1)

# define inverse of logistic function
invlogit <- function(x) exp(x) / (1 + exp(x))

# obtain probability of treatment and whether treatment was observed
p_i <- invlogit(alpha_0 + alpha_1 * x_1 + alpha_2 * x_2 + 0 * rnorm(n, 0, sigma_t^2))
t_i <- p_i > runif(n)

hist(p_i)
summary(t_i)

y_i <- beta_0 + beta_1i * t_i + beta_2 * x_2 + beta_3 * x_3 + rnorm(n, 0, sigma_y^2)

hist(y_i)
summary(lm(y_i ~ t_i + x_2 + x_3))
mean(y_i[t_i == 1]) - mean(y_i[t_i == 0])

summary(x_2[t_i == 1])
summary(x_2[t_i == 0])
cov(t_i, x_2)
```
