---
title: "Proportion Treated"
author: "Waveley Qiu"
date: "2022-02-11"
output: html_document
---

```{r setup, include=FALSE}
# load necessary packages
library(tidyverse)
library(simstudy)
library(MatchIt)
library(dplyr)

# set knitr defaults
knitr::opts_chunk$set(
    echo      = TRUE
  , fig.align = "center"
  , fig.width = 6
  , fig.asp   = .6
  , out.width = "90%"
)

# set theme defaults
theme_set(
  theme_bw() +
  theme(
    legend.position = "bottom"
    , plot.title    = element_text(hjust = 0.5)
    , plot.subtitle = element_text(hjust = 0.5)
    , plot.caption  = element_text(hjust = 0.0)
  )
)

# set color scale defaults
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill   = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete   = scale_fill_viridis_d
```


  \[ h(x) = \log \left( \frac{x}{1 - x} \right). \]

Fix $n \in \mathbb{N}$, and consider a population of individuals $i \in \{ 1, \ldots, n \}$. Let $X_{1i}, X_{2i}, X_{3i} \sim N(0, 1)$ be covariates associated with individual $i$. Let $T_i$ be a binary variable indicating some treatment, and let the data-generating process for $T_i$ satisfy

  \[
    h \Big( P \big(T_i = 1 \; | \; \mathbf{X}_i, \boldsymbol{\alpha} \big) \Big)
    = \alpha_0 + \alpha_1 X_{1i} + \alpha_2 X_{2i} + \varepsilon_i
  \]

for $\varepsilon_i \sim N(0, \sigma_\varepsilon^2)$. That is, the probability of receiving treatment $T_i$ for individual $i$ is determined by $X_{1i}, X_{2i},$ and random noise. 

If we want to set a certain proportion of the population to be treated, we want to fix $E(T_i) = \pi_0$, for a given $\pi_0$. 

$$
\begin{aligned}
E(T_i) &= E[E(T_i \; | \; \mathbf{X}_i, \boldsymbol{\alpha})]
\\
&= E\left[\frac{\exp(\alpha_0 + \alpha_1 X_{1i} + \alpha_2 X_{2i} + \varepsilon_i)}{1+\exp(\alpha_0 + \alpha_1 X_{1i}+ \alpha_2 X_{2i} + \varepsilon_i)}\right]
\\
&= \frac{\exp(\alpha_0 + \alpha_1 E(X_{1i}) + \alpha_2 E(X_{2i}) + E(\varepsilon_i))}{1+\exp(\alpha_0 + \alpha_1 E(X_{1i}) + \alpha_2 E(X_{2i}) + E(\varepsilon_i))}
\\
&= \frac{\exp(\alpha_0)}{1+\exp(\alpha_0)}
\\
\implies \pi_0 &= \frac{\exp(\alpha_0)}{1+\exp(\alpha_0)}
\\
\implies \alpha_0 &= \frac{\log{\pi_0}}{\log(1-\pi_0)}
\end{aligned}
$$

## Treatment Assignment

```{r assign_treatments}
# assign_treatments() produces a vector of treatments.
# inputs:
# - df: input dataframe
# - p: desired proportion of treated individuals
# - alpha_vec: vector of alpha coefficients excluding alpha_0 
# - var_mat: selected variables (also need to specify error term)
# - cov_expval_vec: vector of covariate expected values (if X_i ~ N(0,1), all 0)

assign_treatments <- function(df, p, alpha_vec = c(0.4, 0.5), var_mat = c("L1", "L2"), cov_expval_vec = c(0, 0)){
  
  # turn alpha coefficients into vector
  alpha_vec_loc <- alpha_vec %>% as.vector()
  
  # turn covariates from input df into matrix
  var_mat_loc <- df %>% select(var_mat) %>% as.matrix()
  
  
  
  # turn covariate expected values into vector
  cov_expval_vec_loc <- cov_expval_vec %>% as.vector()
  
  
  # calculating logit parameter
  # a linear combination of:
  # - alpha coefficients
  # - covariate expected values
  # - desired proportion of treated individuals
  
  alpha_0_logit_param <- sum(alpha_vec_loc %*% t(cov_expval_vec_loc)) + p
  
  # sanity check
  # print(alpha_0_logit_param)
  
  # calculate alpha_0 (if E(X_i) all 0, alpha_0 is just simple logit(p))
  alpha_0 <- log(alpha_0_logit_param/(1 - alpha_0_logit_param))
  
  # set treatments
  A_logit <- alpha_0 + var_mat_loc %*% alpha_vec_loc # adding alpha_0 to product of var_mat and alpha_vec
  p_A     <- exp(A_logit)/(1 + exp(A_logit)) # probability is inverse logit of A_logit
  A       <- rbinom(N, 1, p_A) # assigning treatments
  
  # return treatment column
  return(A)
}

```
