---
title: "PS Bootstrap Binary Standardization"
author: "Hun & Amy"
date: "2/9/2022"
output: pdf_document
---


```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
library(simstudy)
library(MatchIt)
library(dplyr)
```

The goal of this document is to combine a lot of the functions that Hun has made and made them into functions that can easily to called for the main loop of the simulations function. 

## Function to find the intercept to attain the desired treatment rate 
```{r message=FALSE, warning=FALSE}
treatment_rate_coef <- function(desired_prop, seq_try, df, logit_formula) {
  # setting up data setting variables 
  seq_select = tibble(seq_try, mean_A = rep(0, length(seq_try))) 
  
  # for loop that goes through all intercept options. 
  for(i in c(1:length(seq_try))) {
    beta_0_treat = seq_try[i]

    # this will be where the beta is updated. Must have beta_0_treat in formula
    A_logit <- eval(parse(text=logit_formula)) 
    p_A     <- exp(A_logit)/(1+ exp(A_logit))
    A       <- rbinom(N, 1, p_A) # this decides treat and and non treated
    
    # getting the mean treated
    seq_select$mean_A[i] = mean(A)
  }
  
  # # plotting visualizing check
  # seq_select %>%
  #   ggplot(aes(x=seq_try, y=mean_A))+
  #   geom_point()
  
  # selecting coeff that gives desired treated untreated proportion 
  if (sum(seq_select$mean_A == desired_prop) > 0) {
    beta_0_treat = seq_select %>%  filter(mean_A == desired_prop ) %>% pull(seq_try)
  } else {
    beta_0_treat = seq_select %>% 
      filter(mean_A < desired_prop + 0.01) %>%
      filter(mean_A > desired_prop - 0.01) %>%
      summarise(
        avg = mean(seq_try)
      ) %>% pull(avg)
  }
  
  # returning the beta value 
  return(mean(beta_0_treat))
  
}

```

## Finding the right beta_effect estimate for binary data  
```{r message=TRUE, warning=TRUE}

binary_outcome_rate <- function(desired_diff, seq_try, df, logit_formula ) {
  seq_select = tibble(seq_try, prop_save = rep(0, length(seq_try)))
  for(i in c(1:length(seq_try))) {
    beta_effect = seq_try[i]
    
    # binary outcome 
    y_logit   <-eval(parse(text=logit_formula ))  
    p_outcome <- exp(y_logit)/(1+ exp(y_logit))
    y_binary  <- rbinom(N, 1, p_outcome)
    
    # finding difference between groups
    prop = tibble(y_binary, A = df$A) %>%
      group_by(A) %>% 
      summarise( mean_out = mean(y_binary)) %>% 
      pivot_wider(
        names_from = A,
        values_from = mean_out
      ) %>% 
      mutate (diff = `1`-`0`) %>% 
      pull(diff)
    
    # having the difference 
    seq_select$prop_save[i] = prop
  }

  # graphing 
  # seq_select %>%
  #   ggplot(aes(x=seq_try, y=prop_save))+
  #   geom_point()

 # saving the best beta. 
  beta_effect = seq_select %>% 
    filter(prop_save < desired_diff + 0.01) %>%
    filter(prop_save > desired_diff - 0.01) %>%
    summarise(
      avg = mean(seq_try)
    ) %>% pull(avg)
 return( beta_effect)
}

# logit_formula <- "beta_effect*df$A + 0.5*df$L2 + 0.3*df$L3"
# seq_try = seq(-4,4, 0.01)
# desired_diff = 0.4
# beta_effect = binary_outcome_rate(desired_diff, seq_try, df, logit_formula)
```


## Generating data 
```{r message=FALSE, warning=FALSE}

data_gen <- function(N, desired_prop) {

  # Setting the Observed Variables 
  L1 <- rnorm(N, 0, 1)
  L2 <- rnorm(N, 0, 1)
  L3 <- rnorm(N, 0, 1)
  
  # adding the covaraites to dataset 
  df <- tibble(L1, L2, L3)
  
  desired_prop = 0.2 # TM - is this correct as a constant or should it be a variable?
  seq_try      = seq(-2,2, 0.01)
  logit_formula <- "beta_0_treat+ 0.5*df$L1 + 0.4*df$L2" 
  
  # get the desited treatment rate 
  beta_0_treat = treatment_rate_coef(desired_prop, seq_try, df, logit_formula)
  
  # Creating the treatment assigment
  A_logit <- beta_0_treat+ 0.5*df$L1 + 0.4*df$L2
  p_A     <- exp(A_logit)/(1+ exp(A_logit))
  A       <- rbinom(N, 1, p_A)
  
  # adding the treated untreated data to the dataset 
  df$A = A
  
  # continuous outcome 
  y_continous <- 1*df$A +  0.5*df$L2 + 0.3*df$L3 +  rnorm(N, 0,1)
  
  df$y_continous = y_continous
  
  
  # binary outcome 
  logit_formula <- "beta_effect*df$A + 0.5*df$L2 + 0.3*df$L3"
  seq_try = seq(-4,4, 0.01)
  desired_diff = 0.4
  beta_effect = binary_outcome_rate(desired_diff, seq_try, df, logit_formula)
  
  y_logit   <-eval(parse(text=logit_formula ))  
  p_outcome <- exp(y_logit)/(1+ exp(y_logit))
  y_binary  <- rbinom(N, 1, p_outcome)
  
  df$y_binary = y_binary
  
  # propensity scores 
  exposureModel <- glm(A ~ L1 + L2 + L3, data = df, family = "binomial")
  #Note that Pr[A=0|L] = 1-Pr[A=1|L]

  #getting estimated propensity score
  df$ps <- predict(exposureModel, df, type = "response") 
  
  return(df)
}

# calling the data gen function 
N = 500
desired_prop = 0.2
df = data_gen(N, desired_prop)
```



## Nearest neighbor propensity score matching
```{r, warning = FALSE}
matched <- matchit(A ~ L1 + L2 + L3, data = df, 
                     distance = "glm", link = "logit",
                     method = "nearest", ratio = 1)

matched_df <- match.data(matched)
```



## simple bootstrap binary 
```{r}
nboot <- 100
simple_bootstap_binary <- function(nboot, matched_df) {

  # set up a matrix to store results
  boots <- data.frame(i = 1:nboot,
                      se_ATE = NA,
                      se_OR = NA,
                      log_OR = NA,
                      mean1 = NA,
                      mean0 = NA,
                      difference = NA
                      )
  # loop to perform the bootstrapping
  for (i in 1:nboot) {
    # sample with replacement
    sampl <- matched_df %>% filter(subclass %in% sample(levels(subclass),500, replace =  TRUE))
    
    bootmod <- glm(y_binary ~ A + ps, data = sampl, 
                   weights = weights, family = binomial)
    
    # create new data sets
    sampl.treated   <- sampl %>% mutate(A = 1)
    sampl.untreated <- sampl %>% mutate(A = 0)
    
    # predict values
    sampl.treated$pred.y   <- predict(bootmod, sampl.treated, type = "response")
    sampl.untreated$pred.y <-  predict(bootmod, sampl.untreated, type = "response")
  
    
    # output results
    boots[i, "log_OR"] <- summary(bootmod)$coeff[2,1]
    boots[i, "se_OR"]  <- summary(bootmod)$coeff[2,2]
    boots[i, "se_ATE"] <- 
      sqrt((summary(bootmod)$coeff[2,2]*mean(sampl.treated$pred.y) *
         (1 - mean(sampl.treated$pred.y)))^2 +
      (summary(bootmod)$coeff[2,2]*mean(sampl.untreated$pred.y) *
         (1 - mean(sampl.untreated$pred.y)))^2)
    
    boots[i, "mean1"]      <- mean(sampl.treated$pred.y)
    boots[i, "mean0"]      <- mean(sampl.untreated$pred.y)
    boots[i, "difference"] <- boots[i, "mean1"] - boots[i, "mean0"]
    
    
  }
  
    mean_log_OR         <- mean(boots$log_OR)
    Empirical_se_ATE    <- sd(boots$difference)
    mean_se_ATE         <- mean(boots$se_ATE)
    Empirical_se_log_OR <- sd(boots$log_OR)
    mean_se_log_OR      <- mean(boots$se_OR)
    ATE                 <- mean(boots$difference)
    lower_CI_ATE        <- ATE - 1.96*Empirical_se_ATE
    upper_CI_ATE        <- ATE + 1.96*Empirical_se_ATE
    lower_CI_log_OR     <- mean_log_OR - 1.96*mean_se_log_OR
    upper_CI_log_OR     <- mean_log_OR + 1.96*mean_se_log_OR
    
    return(tibble(mean_log_OR, Empirical_se_ATE, mean_se_ATE,
           Empirical_se_log_OR, mean_se_log_OR, ATE,
           lower_CI_ATE, upper_CI_ATE, lower_CI_log_OR,
           upper_CI_log_OR))
    
}

simple_bootstap_binary(100, matched_df)

```


## simple bootstrap continous 
```{r}
nboot <- 100
simple_bootstap_continous <- function(nboot, matched_df) {
  # set up a matrix to store results
  boots <- data.frame(i = 1:nboot,
                      se = NA,
                      mean1 = NA,
                      mean0 = NA,
                      beta1 = NA, 
                      ATE = NA
                      )
  # loop to perform the bootstrapping
    
  for(i in 1:nboot) {
    # sample with replacement
    sampl <- matched_df %>% filter(subclass %in% sample(levels(subclass),500, replace =  TRUE))
    
    bootmod <- glm(y_continous ~ A + ps, data = sampl, 
                   weights = weights)
    
    # create new data sets
    sampl.treated <- sampl %>%
      mutate(A = 1)
    
    sampl.untreated <- sampl %>%
      mutate(A = 0)
    
    # predict values
    sampl.treated$pred.y <- 
      predict(bootmod, sampl.treated)
    
    sampl.untreated$pred.y <- 
      predict(bootmod, sampl.untreated)
    
     # output results
    boots[i, "beta1"] <- summary(bootmod)$coeff[2,1]
    boots[i, "se"] <- summary(bootmod)$coeff[2,2]
    boots[i, "mean1"] <- mean(sampl.treated$pred.y)
    boots[i, "mean0"] <- mean(sampl.untreated$pred.y)
    boots[i, "ATE"] <- boots[i, "mean1"] - boots[i, "mean0"]
  }

  Empirical_sd  <- sd(boots$ATE)
  ATE           <- mean(boots$ATE)
  mean_se       <- mean(boots$se)
  lower_CI_ATE  <- ATE - 1.96*mean_se
  upper_CI_ATE  <- ATE + 1.96*mean_se
  ATE_from_beta <- mean(boots$beta1)

  return(tibble(Empirical_sd,ATE, mean_se, lower_CI_ATE,
                upper_CI_ATE, ATE_from_beta))
}

simple_bootstap_continous(nboot, matched_df)
```









